---
title: "Deep Learning MNIST with Keras"
output: pdf_document
---

This is an example from R studio R Keras website.
https://keras.rstudio.com/articles/examples/mnist_cnn.html

# Package installation
```{r}
install.packages("keras")
install.packages("tensorflow")

# To install both Keras library & TensorFlow backend
library(keras)
library(tensorflow)
#Sys.setenv(RETICULATE_PYTHON="/opt/anaconda3/bin/python3")
#install_tensorflow()
```


# The data, shuffled and split between train and test sets
```{r}
mnist   <- dataset_mnist()
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test  <- mnist$test$x
y_test  <- mnist$test$y
```

# tensor
```{r}
dim(x_train)

# visualize digit
digit <- x_train[5,,]
plot(as.raster(digit, max =255))
```

# Redefine  dimension of train/test inputs
```{r}
x_train <- array_reshape(x_train, c(60000, 28 * 28))
x_test <- array_reshape(x_test, c(10000, 28 * 28))
```

# Transform RGB values into [0,1] range
```{r}
x_train <- x_train / 255
x_test  <- x_test / 255
cat('x_train_shape:', dim(x_train), '\n')
cat(nrow(x_train), 'train samples\n')
cat(nrow(x_test), 'test samples\n')
```

```{r}
# Convert class vectors to binary class matrices: one-hot encode
num_classes <- 10
y_train <- to_categorical(y_train, num_classes)
y_test <- to_categorical(y_test, num_classes)
```

## Feed-forward NN
# Define Model -----------------------------------------------------------

```{r}
model <- keras_model_sequential() 
model %>% 
  layer_dense(units = 256, activation = "relu", input_shape = c(28 * 28)) %>% 
  layer_dropout(rate = 0.4) %>% 
  layer_dense(units = 128, activation = "relu") %>%
  layer_dropout(rate = 0.3) %>%
  layer_dense(units = 10, activation = "softmax")
```

```{r}
model %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_rmsprop(),
  metrics = c("accuracy")
)
```

```{r}
history <- model %>% fit(
  x_train, y_train, 
  epochs = 20, batch_size = 128, 
  validation_split = 0.2
)
```


```{r}
plot(history)
```



```{r}
model %>% evaluate(x_test, y_test,verbose = 0)

```

```{r}
model %>% predict_classes(x_test)

```



## CNN

# The data, shuffled and split between train and test sets
```{r}
mnist   <- dataset_mnist()
x_train <- mnist$train$x
y_train <- mnist$train$y
x_test  <- mnist$test$x
y_test  <- mnist$test$y
```

# Redefine  dimension of train/test inputs
```{r}
x_train <- array_reshape(x_train, c(nrow(x_train), 28, 28, 1))
x_test  <- array_reshape(x_test, c(nrow(x_test), 28, 28, 1))
input_shape <- c(28, 28, 1)
```

# Transform RGB values into [0,1] range
```{r}
x_train <- x_train / 255
x_test  <- x_test / 255
cat('x_train_shape:', dim(x_train), '\n')
cat(nrow(x_train), 'train samples\n')
cat(nrow(x_test), 'test samples\n')
```

```{r}
# Convert class vectors to binary class matrices
y_train <- to_categorical(y_train, 10)
y_test <- to_categorical(y_test, 10)
```

# Define Model -----------------------------------------------------------

# Define model
```{r}
model <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, 
                kernel_size = c(3,3), 
                activation = 'relu',
                input_shape = input_shape) %>% 
  layer_conv_2d(filters = 64, 
                kernel_size = c(3,3), 
                activation = 'relu') %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_dropout(rate = 0.25) %>% 
  layer_flatten() %>% 
  layer_dense(units = 128, activation = 'relu') %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_dense(units = 10, activation = 'softmax')
```

# alternative
```{r,eval=FALSE}
model <- keras_model_sequential() %>%
  layer_conv_2d(filters = 32, strides=c(1L, 1L),
                kernel_size = c(5,5), 
                activation = 'relu',
                input_shape = input_shape) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2), strides=c(2L, 2L)) %>% 
  layer_conv_2d(filters = 64, 
                kernel_size = c(5,5), 
                activation = 'relu') %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_flatten() %>% 
  layer_dense(units = 1000, activation = 'relu') %>% 
  layer_dense(units = 10, activation = 'softmax')
```
# Compile model

```{r}
model %>% compile(
  loss = loss_categorical_crossentropy,
  optimizer = optimizer_adadelta(),
  metrics = c('accuracy')
)
```

# Train model

```{r}
model %>% fit(
  x_train, y_train,
  batch_size = 128,
  epochs = 4,
  validation_split = 0.2
)
```

# Evaluate the model

```{r}
scores <- model %>% evaluate(
  x_test, y_test, verbose = 0
)
```

# Output metrics

```{r}
cat('Test loss:', scores[[1]], '\n')
cat('Test accuracy:', scores[[2]], '\n')
```
